### 2018-07-23 articles/cycle/response-referee2.tex

% response-editor.tex
% ----------------------------------------------------------------------
\input{response-header}

\title{Authors' response to Anonymous Referee \#2}
\author{J.~Seguinot, on behalf of all authors.}
%\date{}

\begin{document}
\maketitle
\bigskip

% ----------------------------------------------------------------------
% Interactive comment text begins
% ----------------------------------------------------------------------

\newcommand{\sechead}[1]{\bigskip\noindent\textbf{#1}}
\newcommand{\referee}[1]{\bigskip\noindent\textcolor{darkblue}{#1}}
\newcommand{\msquote}[1]{\begin{quote}\textit{#1}\end{quote}}
\newcommand{\doi}[1]{doi:\allowbreak\href{http://dx.doi.org/#1}{#1}}

Dear Anonymous Referee \#2,

Thank you very much for your detailed review of our manuscript.

% ----------------------------------------------------------------------

\sechead{General comments}

    \referee{%
        This manuscript describes a model study with the Parallel Ice Sheet
        Model applied to the last glacier cycle in the Alps. The climate
        forcing is derived from present day climate of WorldClim and the
        ERA-Interim reanalysis and time-dependent temperature offsets derived
        from the Greenland Ice core (GRIP), from Antarctic ice core (EPICA) and
        Marine sediment core from the Iberian margin (MD01-2444). The study is
        split in two, in the first part analysis of six model simulations (with
        and without precipitation scaling) made on a 2 km resolution grid is
        presented and concluded that out of the three climate forcing records
        used, the EPICA record gives the most realistic ice volume history
        during MIS 4 and 2. The second part analyses simulation made with the
        EPICA forcing record on a 1km grid. The authors draw conclusions about
        the ice cover, ice flow pattern, ice thickness, LGM ice extent, which
        in their model is a transient stage with varying timing across their
        model domain due to glacier dynamics.}

    Thank you very much for this fully accurate summary.

    \referee{%
        The manuscript is well organized
        and clearly written, the missing thing in this study is a discussion of
        and preferably a sensitivity study of the ice dynamic-model assumptions
        made. Is the sliding of the ice age ice sheet realistically modelled
        with pseudo-plastic assumption using the Shallow Shelf approximation?
        How sensitive are the results to the selected model parameters? It is
        briefly mentioned once on page 19, line 8, but a thorough analysis of
        the model sensitivity would strengthen the paper.}

    Thank you.
    We agree that a sensitivity study to ice physics, in particular basal
    sliding, is one of many improvements needed upon our results. However, for
    the kind of exercise presented here, the sensitivity to palaeo-climate
    history, which is virtually unknown and has many degrees of freedom, far
    outweight the model sensitivity to ice physics. From the perspective of
    glacial geology, a 5\,K change in air temperature is more influential than
    an order-of-magnitude increase in ice velocities (perhaps with the
    exception of our conclusions on trimlines, see below). In its current
    state, our manuscript studies the model sensitivity to the primary source
    of uncertainty, which is climate.

    Because the simulations presented here depend on computing
    resources only available through a peer-review application process, we are
    currently unable to perform an additional sensitivity study on basal
    sliding for the entire Alps. However, a regional study on the model
    sentivity to basal sliding is actually
    the topic for a follow-up study currently under preparation for the Rhine
    Glacier \citep{Imhof.etal.2017}, which has a relatively isolated catchment
    for its size as compared to other Alpine Glaciers.

    Meanwhile, the following text was added in the section on trimlines, which
    is certainly the most sensitive aspect of our results to basal sliding:

    \msquote{%
        The remaining discrepancies may relate to temporal migrations of the
        basal thermal boundary, an absence of sliding in warm-based areas, and
        levelling of small-scale topographic features in the 1\,km horizontal
        grid. They call for more detailed comparisons spanning the entire
        Alpine range and specific sensitivity studies to relevant basal sliding
        and ice rheological parameters, and to the uncertain subglacial
        topography.}

    In addition, basal sliding and other sources of uncertainties in ice
    physics where also mentioned in the conclusions:

    \msquote{%
        However, these results are limited by uncertainties on ice physics. The
        till deformation model used here does not hold for sliding over bedrock
        surfaces. On the other hand, the constant friction angle used is
        representative of wet till but weaker basal conditions may have applied
        over saturated lake sediments where they occured. In the absence of ice
        deformation measuremets, a constant rheology was used for temperature
        ice containing more than 1\,\% of liquid water.}

% ----------------------------------------------------------------------

\sechead{Specific comments}

    \referee{%
        I find missing something that indicates that the times are before
        present, as in line 5, line 8 and line 16 on page 1 -- and elsewhere in
        the paper it is written (120-0 ka) should you add before present or BP
        to indicate the time interval?}

    We have now defined this better in the introduction:

    \msquote{
        During the last 800\,ka (thousand years before present)}

    \referee{%
        Could the reason for too large ice volumes when using the GRIP record,
        as mentioned in lines 13-15 on page 9 be due to Arctic Amplification?
        Could that have an effect then as it has now? This is also mentioned in
        line 11 on page 10.}

    These passages refer to the Younger Dryas and Dansgaard-Oeschger events.
    One must keep in mind that a linear scaling was applied to the GRIP record,
    so that it is already ``de-amplified'' before being used as temperature
    forcing. But still, Arctic amplification could be an explanation if it was stronger
    then than for changes that lead the LGM climate. Although we lack expertise
    on this topic, we could not find literature supporting this idea. Instead,
    current theories for the Younger Dryas and Dansgaard-Oeschger events mostly
    involve large amounts of freshwater discharge into the North Atlantic
    implying a more regional mechanism than Arctic amplification. To avoid
    restrictive interpretations, we thus decided not to mention Arctic
    amplification as a possible explanation.

% ----------------------------------------------------------------------

\sechead{Minor comments}

    \referee{\textbf{p.~1, l.~24:}
        suggest: ``have extended well outside their current margins''.}

    Thank you. Done.

    \referee{\textbf{p.~2, l.~22:}
        could you add a reference and a timing for LGM?}

    Done.

    \referee{\textbf{p.~3, l.~1:}
        suggest to replace ``thus'' with ``still''.}

    Done.

    \referee{\textbf{p.~3, l.~3:}
        add s to responses.}

    Done.

    \referee{\textbf{p.~3, l.~18:}
        something missing in the sentence, suggest ``formulation'' after
        ``creep''.}

    Done.

    \referee{\textbf{p.~4, l.~4:}
        suggest to replace ``field'' with ``sheet''.}

    Done.

    \referee{\textbf{p.~5, Fig.~1:}
        c) can you add a scale and maybe indication with a box in b) where this
        extract is from?  ``from the estimate'' (not plural in line 4 of
        caption), Line 6 (PDD) is not acronym for surface mass balance, some
        more explanation is needed here, indicate also, like in the other
        figures that h) is January and i) is July figures.}

    We have added a scale, marked the inset location, and clarified the text.

    \referee{\textbf{p.~6, l.~6:}
        suggest to replace ``of'' with ``with''.}

    Done.

    \referee{\textbf{p.~6, l.~14:}
        suggest ``The climate forcing driving the ice sheet simulations consist
        spatially of a present-day monthly mean climatology...``}

    Done.

    \referee{\textbf{p.~6, l.~15:}
        suggest to delete ``s'' on corrections.}

    Done.

    \referee{\textbf{p.~6, l.~19:}
        add ``mean'' after monthly.}

    Done.

    \referee{\textbf{p.~6, l.~22:}
        note, if true (clarify in Figure caption, see comment above) then the
        reference should be to Fig. 1 i) for summer precipitation.}

    This is true, done.

    \referee{\textbf{p.~6, l.~27:}
        replace ``is'' with ``are''.}

    Done.

    \referee{\textbf{p.~6, footnote:}
        add a reference for the correct formula for the rigidity and clarify
        (add something like, the consequence of this error is that the
        simulations effectively use....) Delete ``in'' before ``a small'' and
        add a quantification of the small change in the length scale – how
        small, is it a few percentage?}

    The correct formula can be found in the study by \citet[p.~443]{Love.1906}.
    The lenght scale of bedrock deformation can be computed as
    ${\alpha=\sqrt[4]{\frac{4D}{(\rho_m-\rho_l)g}}}$ \citep{Walcott.1970}, thus
    the error in lenght scale can be expressed as $\sqrt[4]{1+\nu}-1 = 5.7\%$.
    The footnote was clarified and references were added.

    \referee{\textbf{p.~7, l.~2:}
        ``shipped with WorldClim'' is not clear, please edit, also suggest not
        to use b for topography, s surf and s bed , or h would be better.}

    We clarified the sentence. However, we would like to keep the current
    notation to remain consistent with previous papers \citep{Seguinot.2014,
    Seguinot.etal.2014}.

    \referee{\textbf{p.~7, l.~19:}
        add ``the'' before ``oxygen''.}

    Done.

    \referee{\textbf{p.~7, l.~24:}
        ``and within a rectangular region ..'' is not clear, edit this text.}

    The sentence was reorganised.

    \referee{\textbf{p.~8:}
        add ``acceleration of'' before gravity, add a reference for the ideal
        gas constant?}

    Done.

    \referee{\textbf{p.~9, l.~14--15:}
        see comment above, could this be an example of Arctic amplification?}

    Please refer to my comment above.

    \referee{\textbf{p.~9, l.~19:}
        add ``a'' before ``very'' and suggest to turn the sentence around, the
        EPICA simulations are in a good agreement with the data.}

    Done.

    \referee{\textbf{p.~9, l.~20:}
        followed by first a retreat and then a standstill? replace ``blue''
        with ``red''.}

    Done.

    \referee{\textbf{p.~9, l.~21:}
        suggest : The simulations forced by the GRIP palaeo-temperature forcing
        yield ... (blue curves)''.}

    Done.

    \referee{\textbf{p.~10, l.~2:}
        suggest to add ``followed by a rapid retreat''.}

    Done.

    \referee{\textbf{p.~10, l.~3:}
        why state ``two or three'' in intro the suggestion is either 4 major or
        15, is there a reference for 2 or 3?}

    The numbers 4 and 15 relate to the total (minimum) number of Pleistocene
    glaciations. However, evidence for pre-LGM glaciations during the last
    glacial cycle is debated and very limited (see later part of the
    introduction). We added the word ``Pleistocene'' in the introduction and
    clarified the concerned sentence about model results.

    \referee{\textbf{p.~10, l.~10:}
        suggest to replace ``lower'' with ``smaller''.}

    Done.

    \referee{\textbf{p.~10, l.~11:}
        is this due to Arctic Amplification?}

    Please refer to my comment above.

    \referee{\textbf{p.~10, l.~12:}
        suggest to replace ``least'' with ``the smallest''.}

    Done.

    \referee{\textbf{p.~11, Fig.~3 caption:}
        what does ``cumulative'' mean here? Do you mean maximum in each
        location?  clarify what the black line indicates. No solid red line is
        visible in figures. What is meant with ``reasonable'', maybe replace
        with ``realistic''? suggest to replace ``cover'' with ``extent''.}

    We replaced ``cumulative extent of modelled ice cover'' with ``modelled
    maximum extent'', ``reasonable'' with ``realistic'', and ``red line'' with
    ``black line'', which is what we meant.

    \referee{\textbf{p.~11, l.~2:}
        suggest to replace ``cumulative'' with ``maximum extent in each area''
        or something similar.}

    As for the caption, we replaced with ``modelled maximum extent''.

    \referee{\textbf{p.~11, l.~1--3:}
        how sensitive is the model to different parameters in the applied
        sliding formulation?}

    Please refer to my comment above.

    \referee{\textbf{p.~11, l.~4--5:}
        ``outside this benchmark'' clarify if you mean spatially or temporally.}

    Done. We mean spatially.

    \referee{\textbf{p.~11, l.~6--11:}
        is this text better fitted in a discussion section?}

    We have included this discussion here as we think that understanding the
    simiarities and differences in the spatial patterns of glaciation between
    the different palaeo-temperature records used is necessary to justifying
    the choice of the EPICA forcing used in the next section. We think that
    this non-standard outline is the best way to present all results in a
    single manuscript.

    \referee{\textbf{p.~12, l.~2:}
        suggest to replace ``higher'' with ``larger''.}

    Done.

    \referee{\textbf{p.~12, l.~4:}
        do you mean to refer to Fig. 3?}

    No, but the reference to Fig.~2 was misplaced. This has been corrected.

    \referee{\textbf{p.~12, l.~29--32:}
        does this text fit better in method section?}

    We agree that this text does not really fit in a results and discussion
    section. However, this information is specific to Sect.~4. We feel that
    moving these and the corresponding sentences in Sect.~3 to the methods
    section would increase the complexity of the manuscript.

    \referee{\textbf{p.~13, Fig.~4 caption:}
        the 200 m surface contours are not clearly visible in the figure, can
        they be made sharper, or just skipped? Line 3, something like ``are
        shown'' is missing. Suggest to replace``background'' with ``bedrock''.
        This is the first time ``Natural Earth Data'' is mentioned, should that
        be in the section on the data? Suggest to replace ``Gray fields'' with
        ``shaded gray area'' and ``boundaries'' with ``timing''.}

    The 200\,m contours were thickened, the caption was reworked as suggested,
    and the reference to Natural Earth Data was moved to the caption of Fig.~1.

    \referee{\textbf{p.~15, l.~8:}
        replace ``was'' with ``is''.}

    Done.

    \referee{\textbf{p.~15, l.~25:}
        ``occurred''.}

    Done.

    \referee{\textbf{p.~16, Fig.~5:}
        in figure the color bar is written to indicated maximum surface
        elevation, but in figure caption the maximum ice thickness, which is
        correct? The surface elevation contours are not clearly visible. The
        dark orange color in the figure, that covers the central part of the
        ice sheet is not (?) in the bar on the left (or is it before 27 ka
        BP?).}

    The map shows maximum thickness. The colourbar label was corrected.
    Surface elevation contours were thickened. Indeed, in much of the mountain
    area maximum thickness is reached before 27\,ka. This is now clarified in
    the figure caption.

    \referee{\textbf{p.~17, l.~30:}
        suggest to replace ``have'' with ``could''.}

    Done.

    \referee{\textbf{p.~18, Fig.~6:}
        the 200 m contours are not clearly visible.}

    The 200\,m contours were thickened.

    \referee{\textbf{p.~19, l.~7:}
        could you add what the modelled regional ice thickness is, for
        comparison?}

    Done.

    \referee{\textbf{p.~20, Fig.~7:}
        add info about the gray areas indicating MIS 4 and 2 ``Isolated patches
        indicate periodic surges from tributary glaciers'' needs more
        explanation and it is not clear what is meant. Does the model simulate
        periodic surges?}

    We added the info about MIS. The wording ``periodic surges'' was badly
    chosen and replaced by a more cautious ``episodic advances''. These
    advances are likely due to thermodynamical feedbacks modulated by
    fluctuations in the climate forcing, but they were not explored in detail
    and perhaps depend on model resolution.

    \referee{\textbf{p.~21, l.~21:}
        here is a reference for 2 or 3 glaciations (see comment above) but in
        the intro is only mentioned 4 or 15, suggest to change text to
        harmonize.}

    See my previous comment about last glacial cycle and older glaciations.

    \referee{\textbf{p.~21, l.~27:}
        suggest to edit ``the study consists of'' or start for example like
        ``In this study the model has been applied...''}

    Done.

    \referee{\textbf{p.~21, l.~28:}
        how important is it that the model has been validated for the
        Cordilleran ice sheet? Will that support the choices of the sliding
        model applied in the Alps? I think that it would be valuable for this
        study to do a sensitivity runs for at least some of the model parameter
        choices.}

    The reference to the Cordilleran ice sheet was removed here. As mentioned
    in the later part of the conclusions, it is clear to us that additional
    sensitivity studies are needed. With this study, we hope to have
    highlighted areas of research were data-model comparison could be pursued
    further. However, this will require not only more model runs, but also a
    more systematic classication of geological data in digital databases.

    \referee{\textbf{p.~22, l.~6:}
        add ``records'' after forcing.}

    We replaced ``forcing'' by ``records''.

    \referee{\textbf{p.~22, l.~19:}
        why do you add ``potentially'' here? Isn't this a firm conclusion from
        you study?}

    The timing of the LGM extent depends on the actual climate history of the
    European Alps, which was almost certainly different and more complex than
    our climate forcing scenarii.  Although our simulations show that even a
    spatially homogeneous forcing produces spatial variations in the timing of
    maximum extent, on can not exclude that a different climate history might
    counteract these variations.

    \referee{\textbf{p.~22, l.~20:}
        suggest to replace ``higher'' with ``larger''.}

    Done.

    \referee{\textbf{p.~22, l.~22:}
        same as above, why ``potentially'' here?}

    We removed the first ``potentially''. Nevertheless, the actual number and
    timing of glacier advances onto the foreland depends on Alpine climate
    history, which may have been very different from the forcing used in
    our simulations.

    \referee{\textbf{p.~22, l.~24:}
        suggest to replace ``nevertheless'' with ``however'' or edit the
        sentence.}

    Done.

    \referee{\textbf{p.~22, l.~26--28:}
        the paper would be able to give stronger conclusions with sensitivity
        study, suggest to edit the sentence by replace ``statements'' with
        ``limitations'' or ``drawbacks'' and ``last glacier cycle ice dynamics
        in the Alps'' is not easy to read.}

    Done. The complicated sentence was rephrased.

    \referee{\textbf{p.~23, l.~2:}
        replace ``mode'' with ``more''.}

    Done.

    \referee{%
        Figures are generally clear and well set up.  The surface contours in
        Figs 4,5 and 6 is not clearly visible in my printout and could maybe
        become clearer?}

    Thank you. The 200\,m contour levels were thickened on all figures. We
    thank you again very much for the time and effort you put into this very
    meticulous review.

% ----------------------------------------------------------------------
% References

\bibliographystyle{copernicus}
\bibliography{../../references/references}

\end{document}