### 2018-07-23 articles/cycle/response-referee3.tex

% response-editor.tex
% ----------------------------------------------------------------------
\input{response-header}

\title{Authors' response to Anonymous Referee \#3}
\author{J.~Seguinot, on behalf of all authors.}
%\date{}

\begin{document}
\maketitle
\bigskip

% ----------------------------------------------------------------------
% Interactive comment text begins
% ----------------------------------------------------------------------

\newcommand{\sechead}[1]{\bigskip\noindent\textbf{#1}}
\newcommand{\referee}[1]{\bigskip\noindent\textcolor{darkblue}{#1}}
\newcommand{\msquote}[1]{\begin{quote}\textit{#1}\end{quote}}
\newcommand{\doi}[1]{doi:\allowbreak\href{http://dx.doi.org/#1}{#1}}

Dear Anonymous Referee \#3,

Thank you very much for your detailed review of our manuscript.

    \referee{%
        This is an excellent manuscript and I highly recommend publication
        after a few minor changes. The authors present a well thought out
        modelling experiment which they combine (albeit in a qualitative
        manner) with extensive palaeo-glaciological data and cumulative work. I
        can see this work being extended into more extensive and rigorous work
        (RCM forcing, ice dynamic sensitivity, quantitative fitting to
        geomorphological record etc), but this is an important leap forward.}

    Thank you very much for these supportive words!

% ----------------------------------------------------------------------

\sechead{Abstract}

    \referee{\textbf{p.~1, l.~2:}
        ``pioneer'' should be ``pioneering''.}

    Thank you. Corrected.

    \referee{\textbf{p.~1, l.~16:}
        I think the finding that you get asynchronous glaciation extents with a
        uniform climate offset is due to glacier hypsometry and setting should
        be stated here. i.e.  that the timing of maximum glaciation and
        recession isn't purely a function of climate. This finding needs to be
        highlighted better in the abstract.}

    Thank you for pointing this out. We have reworked the last sentence in
    the abstract:

    \msquote{%
        Finally, despite the uniform climate forcing, differences in glacier
        catchment hypsometry cause the Last Glacial Maximum advance to be
        modelled as a time-transgressive event, with some glaciers reaching
        their maximum as early as 27\,ka, and others as late as 21\,ka.}

% ----------------------------------------------------------------------

\sechead{Introduction}

    \referee{\textbf{p.~2, l.~28:}
        Ballantyne and Stone (2015) should be added to this list.}

    Thank you. This is relevant indeed. We have added the reference in the
    introduction and in the corresponding section of the discussion.

    \referee{\textbf{p.~2, l.~34:}
        It should be stated that it could be a consequence of both
        glaioclimatic interactions and uncertainties in dating methods.}

    We have added ``or both''.

    \referee{\textbf{p.~3, l.~2:}
        These points serve the literature well to highlight gaps for future
        research. However, I would argue that you do not get very far here on 1
        and 5 and do not completely solve the other 3 points. Your text
        reflects these shortcomings very well, for which you should be
        applauded. Though I think at this stage of the manuscript, your
        statement of intent, you should state that you do not claim to solve
        these questions, but rather push forward on all of them using your new
        approach of ice sheet modelling.}

    Thank you.
    In order to clarify our statement of intent, we added the following text:

    \msquote{%
        Additional geological research will be needed to complete our
        knowledge. But here, we intend to explore these open questions from a
        new angle and [use the Parallel Ice Sheet Model ...].}

% ----------------------------------------------------------------------

\sechead{Section 2.6}

    \referee{\textbf{p.~7:}
        The spatial distribution of your modern climate variables (precip,
        temp) will be massively influenced by elevation. Though there is a
        lapse rate, does this pattern of high precip and low temp over
        mountains remain throughout the simulation despite ice surface
        topography, and if so, how does this influence your results?}

    The temperature lapse rate corrections dampen local high temperatures in
    ice-filled valleys. The precipitation pattern, on the other hand, remains
    the same throughout the simulation, and this results in a higher accumulation
    above the location of present-day mountains even as they are buried by ice
    in the model.

    Although we expect that small-scale variations in surface mass balance are
    smoothed out by ice flow, the presence of an ice sheet may have
    redistributed precipitation on a larger scale, for instance from the inner
    ranges to the foothills, or inequally between north and south of the Alps,
    etc. However, global precipitation reductions appear to have little effect
    on the average ice thickness (Fig.~2b, compare dark and light curves).
    Nevertheless, the new text shortly mentions this caveat:

    \msquote{%
        Besides [circulation changes], changes in the ice surface topography
        may have redistributed orographic precipitation.}

    Our conclusions also highlight the need for a more realistic forcing.

    \msquote{%
        Using [...] a more realistic climate forcing based on regional
        circulation model output [...] future modelling studies will certainly
        be able to quantify uncertainties associated with some of the above
        limitations.}

% ----------------------------------------------------------------------

\sechead{Section 3}

    \referee{\textbf{p.~7, l.~14:}
        You keep mentioning the number of processors. I find this information
        slightly irrelevant, and it will soon become outdated as processing
        speed and models increase (GPUs for example). The only way it will
        serve the community is if there is a full description of the computer
        set up. For example, it could be that the simulations took 4 days on
        144 processors, but the processors were slow. Suggest removing these
        references.}

    We removed mentions of computing times and numbers of processors.

    \referee{\textbf{p.~10, l.~14:}
        This is a more general point. You eventually choose the EPICA record,
        and for justifiable reasons based on comparison to reconstructed ice
        extent and timing.  However, this is likely coincidence. EPICA is
        likely a complex record containing global and local antarctic
        influences upon climate. The ``real'' climate over the alps during
        glaciation is like decoupled from that of Antarctica to an extent.
        Therefore, the match you find is not an inference about climate, as
        different combinations of offsets may have made the same result
        (smoothed GRIP to remove some of the D-O scale noise?). You should make
        this explicit somewhere in the manuscript.}

    We agree. The following text was added in the conclusions:

    \msquote{%
        This suprising result is likely coincidental as there is no direct link
        between European and Antarctic climate. This highlights the need for
        more quantitative reconstructions of European palaeoclimate.}

    \referee{\textbf{p.~11, l.~2:}
        First sentence needs reconsidering as it is slightly broken in its
        current form. Perhaps ``Figure 3a shows the cumulative extent of
        glaciated area during MIS2''.}

    We reworked the first sentence. It now reads:

    \msquote{%
        During MIS 2, all six simulations yield comparative modelled maximum
        extents (Fig.~3a).}

    \referee{\textbf{p.~11, l.~10:}
        Is it possible some ice is missing from the geological reconstruction
        in some instances? I guess some outlets are well constrained, whilst
        others areas could be ``filled in'' by this modelling experiment.}

    Yes, this is possible. Parts of the reconstructed LGM ice limits are
    uncertain due to the lack of moraines or dated evidence. However, without
    very detailed knowledge about palaeoclimate, and tight constraints on ice
    physics, the model won't replace observations. Actually, we are very
    confident that the model-data discrepancies in ice extent discussed in this
    passage are beyond uncertainties.

% ----------------------------------------------------------------------

\sechead{Section 4}

    \referee{\textbf{p.~14, l.~15--31:}
        I find this description of sites and timing of glacier extent compared
        to dates difficult to follow. I suggest a new figure to convey this
        important comparison: Have the reconstructed and modelled ice extents
        at key times for each of the mentioned glaciers on several smaller
        maps, including geochronological constraints.}

    This would be very useful but to our knowledge, no geological data have
    been made available that would allow such a comparison. The review by
    \citet[Fig.~5]{Wirsig.etal.2016} which we refer to is the most up-to date
    compilation of LGM ages in and around the Alps, and deglacial outlines like
    there exists for the Eurasian and Laurentide ice sheets are unfortunately
    not available.

    \referee{\textbf{p.~17, l.~17:}
        This finding is important and should be highlighted better in abstract
        and conclusion.}

    This finding was better highlighted in the abstract (see our previous
    comment). However, we feel it is already well expressed by the following
    bullet point in the conclusion:

    \msquote{%
        The LGM (maximum) extent was a transient stage in which glaciers were
        out of balance with the contemporary climate. Its timing potentially
        varied across the range due to inherent glacier dynamics (Sect. 4.3).}

    \referee{\textbf{p.~17, l.~22:}
        Is the model recreating possible surging? Or would this be an
        over-interpretation given the uncertainty in climate and physics. Seems
        to fit with the enthalpy model of Benn and others for surging. As
        reads, it suggests that these areas were possible palaeo-surges, please
        clarify.}

    Our simulations reproduce a thermodynamical instability akin to a surge.
    The internal energy (so-called enthalpy) conservation scheme used in PISM
    allows for variations in ice temperature and liquid water content
    \citep{Aschwanden.etal.2012}, which may result in thermodynamical
    oscillations \citep{Pelt.Oerlemans.2012, Feldmann.Levermann.2017}.

    Unlike periodic surges \citep[as described by, e.g.,][]{Sevestre.Benn.2015},
    we interpret the model results as triggered by an initial perturbation of
    mass (and energy) caused by rapid climate cooling and inefficient mass
    transfer to lower elevations. We have rephrase the following sentence:

    \msquote{%
        Several Alpine lobes surge and overshoot their balanced extent before
        thinning and receding towards the mountains as they warm towards
        thermodynamical equilibrium.}

    \referee{\textbf{p.~19, l.~9--12:}
        On trimlines: I think you justify well why you haven't yet modelled the
        sensitivity to your trimline result - previous work backs this up. But,
        I think you should consider the following: Adding a plus/minus to your
        results to reflect the uncertainty.  The importance of resolution -
        many trimlines will be below the resolution of your model, so perhaps
        aren't resolved enough for model-data comparison. They may have acted
        to deflect ice flow around mountain peaks for example. This resolution
        caveat should be mentioned. I would be surprised if all trimlines are
        subglacial transitions as this paper suggests - perhaps you need to
        directly challenge the geochronological community to find better
        trimline constraints (sub/supra) as a statement in this paper. Your
        mean value of 861 m is unrepresentative of your sample. Your sample is
        highly skewed, so a modal value (1050 m ish from Fig 6) is more
        appropriate. A similar finding with a similar approach has already been
        found for the British Isles, with the added constraint of GIA
        observations. I suggest referencing Kuchar et al. 2011 for this
        reason.}

    Thank you.
    This is a valid point. However we found an error in our figure: the data
    plotted on the histogram (Fig.~3b) were different from those in the scatter
    plot (Fig.~3a). These data were raw interpolated ice
    thicknesses, i.e. they missed a correction for altitude shifts between our
    initial basal topography and the observed trimline elevations.

    The corrected histogram is closer to symmetry. A slight skewness remains,
    but this is not really a result of our simulations. In fact the
    distribution of observations itself is skewed, with more trimlines observed
    at higher elevations, and there is a tendency for the differences between
    the modelled ice surface and trimlines to increase with altitude.

    The modal value (now ca.~950\,m) corresponds to a cluster of trimline
    locations around 2850\,m in the Aletsch Glacier (south of Jungfrau) region
    where trimlines are well visible in the landscape. We prefer to stick with
    the mean because the number of trimlines mapped is not necessarily
    representative of their lenght, and may depend on other factors such
    as bedrock lithology and post-glacial erosion.

    We now use the standard deviation (197\,m) as a measure of variance, but
    not uncertainty, which would require more sensitivity tests. We also
    mention resolution as one of several caveats in the text:

    \msquote{%
        The remaining discrepancies may relate to temporal migrations of the
        basal thermal boundary, an absence of sliding in warm-based areas, and
        levelling of small-scale topographic features in the 1\,km horizontal
        grid.}

    We added a reference to the study by \citet{Kuchar.etal.2012} although this
    approach is not doable for the Alps.

    \msquote{%
        Unfortunately, validation through the bedrock uplift rate
        \citep[cf.][]{Kuchar.etal.2012} is not doable in the Alps due its lower
        values, active tectonics and uncertainties on geological properties
        \citep[cf.][]{Mey.etal.2016}.}

    \referee{%
        A philosophical but important point is that your discussion throughout
        is written from the standpoint of the geochronological/geomorphological
        data and reconstructions as being ``truth''. It should consider
        somewhere that perhaps data is missing as it is hard-won in the places
        it exists and interpretations may be slightly wrong. The model is also
        not the ``truth'' and there is probably a blurred line in-between upon
        which we can proceed.}

    Thank you. Indeed the quality and availability of geomorphological and
    geochronological data is very variable across the Alps, and long-standing
    interpretations are sometimes proven wrong
    \citep[e.g.,][]{Monegato.etal.2007}. Also, parts of the Alpine glacial
    geology community will certainly disagree with your comment and find our
    challenging of the trimline assumption quite provocative, given that it
    has been around for at least a century \citep{Penck.Bruckner.1909}.

    \referee{\textbf{p.~21, l.~11--19:}
        Be really clear here that these are modelled, and perhaps not
        geologically recorded, advances of the ice sheet. If there is no data,
        it might be correct, might be just a modelled result.}

    Actually we meant geologically recorded asymmetry. Obviously, this needed
    clarification. This sentence now reads:

    \msquote{%
        The observed asymmetric extent of ice north and south of the Alps can
        be explained by the modelled transient nature of the LGM extent without
        involving north-south gradients in temperature and precipitation change
        (Sect.~4.1).}

    Finally, we thank you again very much for the time and effort you put into
    our manuscript.

% ----------------------------------------------------------------------
% References

\bibliographystyle{copernicus}
\bibliography{../../references/references}

\end{document}