### 2018-08-16 articles/cycle/response-editor3.tex
Adress last comments from the editor.

% response-editor.tex
% ----------------------------------------------------------------------
\input{response-header}

\title{Authors' response to the Editor}
\author{J. Seguinot, on behalf of all authors.}
%\date{}

\begin{document}
\thispagestyle{empty}
\maketitle
\bigskip

% ----------------------------------------------------------------------
% Interactive comment text begins
% ----------------------------------------------------------------------

\newcommand{\sechead}[1]{\bigskip\noindent\textbf{#1}}
\newcommand{\referee}[1]{\bigskip\noindent\textcolor{journalname}{#1}}

    Dear Andreas Vieli,

    Thank you very much for supporting our manuscript for publication in
    \emph{The Cryosphere}. We have adressed your comments and performed a few
    last changes on the manuscript as detailed below.

% ----------------------------------------------------------------------

\sechead{Editor comments}

    \referee{\textbf{p.~1, l.~14-17:}
        somewhat odd phrasing: `...cause ...to be modelled...', maybe
        `...produces in the model...' or similar would be better.}

    Thanks. Done.

    \referee{\textbf{p.~3, l.~4:}
        `it still remains incompletely known' is awkward phrasing, maybe better
        say: `...knowledge remains incomplete...'}

    We replaced with ``uncertainties remain on''.

    \referee{\textbf{p.~3, l.~7-8:}
        the added sentence starting with `Additional...' seems slightly out of
        place here (or the flow is someaht disrupted). Maybe better move it to
        end of nextz paragraph (p.~3, l.~14).}

    Done.

    \referee{\textbf{p.~3, l.~9:}
        starting a paragraph with a `But...' is awkward, maybe change back to
        `Here we...' in particular when you move the sentence before.}

    Done.

    \referee{\textbf{p.~3, l.~32:}
        would `different' instead of `distinct' not be more appropriate.}

    We replaced with ``different''.

    \referee{\textbf{p.~4, l.~8:}
        `In the SIA, topographic roughness using a bed smoother range of 5 km
        (Schoof, 2003).' I really do not understand this sentence (what is
        message), nor its connection to the sentence before.}

    Sorry, two words were missing here: ``topographic roughness is parametrized
    using a bed smoother range of 5\,km.'' We have moved this sentence directly
    after the previous mention of the SIA.

    \referee{\textbf{p.~4, l.~24:}
        `...without horizontal transport...'?}

    Done.

    \referee{\textbf{p.~4, l.~25:}
        maybe simplify to `...is assumed to drain instantaneously and is
        removed....'}

    Done.

    \referee{\textbf{p.~7, l.~3:}
        `above average' of what? What is reference?}

    We have changed to ``above annual lapse rates measured in the Alps''. The
    reference describes lapse rate measurements.

    \referee{\textbf{p.~14, fig.~4:}
        are the surface contours really in 200m intervals?, that would make the
        ice surface only a bit over 1200m high. Or do you leave some out?  You
        should add some labels to the surface contour lines so one can judge
        the surface elevation better.}

    An incorrect version of the figure missing 200m contours was uploaded. This
    has been corrected. Thank you very much for noticing.

    \referee{\textbf{p.~13, l.~30 -- p.~14, l.~3:}
        but jouvet (TC) claims (based on tracking erratics) that north south
        contrast in in precip is needed. Is this still consistent with this
        work? Maybe should be included. This is mentioned on next page but
        would be relevant here as well.}

    The reference was added. However, to avoid redundancy with the next section
    on flow patterns, we would prefer to limit the discussion in this section
    to modelled ice marginal positions.

    \referee{\textbf{p.~14, fig.~4 caption:}
        for `(b)' it should say what the vertical black line (time of max
        extent in (a) at 24.57 ka???) refers to.}

    We have made explicit that: ``The black vertical line indicates the
    modelled age of maximum ice cover at 24.57\,ka.''

    \referee{\textbf{p.~15, l.~31:}
        `simulation' (not sumulation)}

    Done.

    \referee{\textbf{p.~16, l.~12:}
        `Fig. 5' has nothing to do with `reviews', what do you mean with `Fig.
        5 for reviews'???}

    With ``cf.~Ivy-Ochs, 2015; Wirsig et al., 2016, Fig. 5 for reviews'', we
    meant to refer to Fig.~5 in the paper by Wirsig et al. (2016). We have
    removed ``Fig.~5'' in this particular instance.

    \referee{\textbf{p.~19, l.~15-19:}
        in this context of sensitivity to basal sliding parameter (and lacking
        exploration) maybe some reference to the Phd-thesis of Becker which
        includes at least some preliminary (mostly steady state) sliding
        investigation could be made.}

    Done.

    \referee{\textbf{p.~21, l.~19:}
        `last major Alpine reliefs' ...? Do you mean the `last foothills' of the
        alps? Clarify.}

    We have replaced this formulation with ``the outermost limestone reliefs''.

    \referee{\textbf{p.~23, l.~1:}
        is it not rather `glacier physics' as the till is not ice (or ice and
        glacier physics).}

    Correct. We have changed to ``glacier physics''.

% ----------------------------------------------------------------------

%\sechead{New references}

    We thank you again very much for your editorial work on our manuscript.

\end{document}