### 2021-05-16 articles/erosion/response-referee1.md
Add review from anonymous referee #1.

<!-- Copyright (c) 2021, Julien Seguinot (juseg.github.io), Ian Delaney
Creative Commons Attribution-ShareAlike 4.0 International License
(CC BY-SA 4.0, http://creativecommons.org/licenses/by-sa/4.0/) -->

# Authors' response to Anonymous Referee #1

Dear Anonymous Referee #1,

Thank you very much for your detailed review of our manuscript.

> The study of Seguinot and Delaney presents a time-integrated model on the
> glacial erosion potential over the last 120 ka in the Alps. Authors tie the
> calculations on a previous study by Seguinot et 2018 who modeled the glacial
> extend over Alps’ last glacial cycle by testing three different records on
> palaeo-temperature (GRIP, EPICA, and MD01-2444). Calculated basal velocities
> which base on the Parallel Ice Sheet Model PISM were used to test different
> glacial postulated erosional laws (Koppes et al., 2015; Herman et al., 2015,
> Humphrey and Raymond, 1994; Cook et al., 2020) focusing on the Rhine glacier
> area.

> The main achievement of the study is to combine a realistic, high-resolution
> model of glacial extension over a glacial cycle in a mountain range with
> assumptions on glacial erosion. Even though the utilized PISM code does not
> account for feedbacks of glacial erosion on bed topography, which makes it
> not directly comparable with common landscape codes on glacial erosion (e.g.
> Egholm et al. 2009; Sternai, et al., 2013), some relevant findings could be
> derived by this integrated high-resolution approach. This includes the
> observation of low erosion during glacier advance and maximum glacier
> extension, and the role of profile steeping during deglaciation and related
> increasing erosion rates.

## General comments

> Authors decided to keep the paper short, referring most information on
> glacial model set up and resulting ice cover to the Seguinot et 2018 paper.
> This makes it no always very easy to read and needs checking in the original
> manuscript (for example, ice cover deviations model – field bases LGM
> extension).

> The introduction gives a nice, relatively long (in relation to other
> chapters) overview on different aspects of glacial erosion and resulting
> features, and guides through problems to investigate them. Here, I don’t see
> very clear how different aspects or parts in the introduction are
> specifically addressed in this study, i.e. which problems are exactly aimed
> to be solved of this large portfolio of mentioned gaps in the understanding
> of glacial processes. I think this could be better formulated and balanced.

> I think the study could also well contribute to the discussion on the
> elevation distribution of the cumulative glacial erosion over several cycles
> in the Alps (high elevations vs low elevation, e.g. Valla et al., 2011). A
> plot showing the integrated glacial erosion potential over (e.g. 100 m)
> elevation bins would easy to do and eventually an enlightening supplement
> summing up Fig. 5a.

> I would also recommend to make some regional statements on the distribution
> of the glacial erosion potential (Fig. 2). It is quite obvious that some
> tectonic massifs can readily be discriminated, e.g. the Tauern Window, Oetzal
> Crystallin complex, the Aare Massif, Mt. Rosa - Gran Paradiso and Pelvoux
> Massif. Is it the steepness of these low erodible crystallin massifs making
> the erosional potential appearing strikingly high? I think this should/could
> be discussed..

## Specific comments

> Line 24 Maybe a bit odd to use 4 very old references from the Alps only (you
> hardly used more than 2 references throughout the MS and there are also high
> variety of glacial landforms in other mt ranges as you mentioned).

> Line 33 I think that’s not very easy to understand what landform you refer
> to? The “periglacial blockfields topped by glacial erratic boulders” (Wirsig
> et al., 2018)? Eventually be more precise on this would help readers.

> Line 101 Could be misleading as (glacial and periglacial) cirque erosion
> processes are not really covered by any glacial erosion law discussed here
> (or any other I am aware of, Sanders et al., 2012).

> Line 116  ..while much OF the intra-montane..

> Line 127 Higher precipitation increases ice flux and thus erosion, I guess?
> Would eventually helpful to mention (even though it might be referred in
> Seguinot et al., 2018).

> Line 140 Can you be more precise what you mean by realistic? You mean because
> of the localized pattern? Maybe I am wrong but shouldn’t the erosion
> potential at least at the lake Constance overdeepening (Fig. 7) not in the
> order of hundreds of meters (cf. Preusser et al., 2010) and the best fit
> rather (b) or (c) – at least from what I read from the transect (e-h)?
> Visually (a)-(d) seem not to correspond to (e)-(h) if (a)-(d) is also
> presented in meters (annotation at the bar is missing). For example, in (b)
> the maximum erosion potential is like >>1000 m (if in meters) while in
> corresponding (f) it seems clearly lower than 1000 m.

> Furthermore, isn’t the impression of the fit potentially very dependent on
> the initial model parametrization (Seguinot et al., 2018), i.e. the ice flux
> velocity? You should address these dependences!

> Line 163 Observing Alpine topography I find this result important, which can
> maybe also serve as explanation why e.g. (low erodible) areas away from the
> big troughs covered during glacial maxima only, do surprisingly often show
> no/very low degree in glacial modification (e.g. Ticino; Kelly et al., 2004).

> Line 187 “time-transgressive radial pattern”. I don’t understand what you
> mean..

> Line 195 Very much share this view!

> Line 197 Would recommend to be more precise. Guess you know that there are
> many, many cirques in the Alps as low as 1500 m (and even below) e.g. forming
> in areas outside the connected ice stream network in the SW and easternmost
> Alps. This can be revealed by a quick check at any higher resolution DEM or
> google earth.

> In the MS the word “yet” is very often used. Eventually consider reducing.
> The frequency is a bit irritating when reading.

## Figures

> Fig. 2 Please indicate the outline of the connected ice stream network during
> the LGM (like in Fig. 4 of Seguinot et al., 2018). Especially in the eastern
> and SW Alps coverage largely deviates from what has been suggested from field
> data compilation (e.g. Ehlers and Gibbard, 2004). Even though outcomes might
> not be changing much, it might be helpful to know how much the %overlap is -
> erosional potential might change as fluvial topography turns into glacial one
> (e.g. Harbor, 1988) and this is probably not what you want to mix, I guess.

> Fig. 5 I have to confess Fig. 5a surprises me, there are really glaciers as
> low as 500 m.a.s.l (and even below) from 110 – 40 ka in the Seguinot et al.,
> 2018 model?

## References not in the manuscript:

> Egholm, D.L., Nielsen, S.B., Pedersen, V.K., Lesemann, J.E., 2009. Glacial
> effects limiting mountain height: Nature, v. 460, p. 884–888.

> Ehlers, J., Gibbard, P.L., 2007. The extent and chronology of Cenozoic Global
> Glaciation. Quat. Int. 164–165, 6–20.

> Harbor, J.M., Hallet, B., and Raymond, C.F., 1988. A numerical model of
> landform development by glacial erosion: Nature, v. 333, p. 347–349.

> Kelly, M.A., Buoncristiani, J.F. and Schlüchter, C., 2004. A reconstruction
> of the last glacial maximum (LGM) ice-surface geometry in the western Swiss
> Alps and contiguous Alpine regions in Italy and France. Eclogae Geol. Helv.
> 97, 57–75.

> Mey, J., Scherler, D., Wickert, A.D., Egholm, D.L., Tesauro, M., Schildgen,
> T.F., and Strecker, M., R., 2016. Glacial isostatic uplift of the European
> Alps: Nature Communications, v. 7.

> Preusser, F., Reitner, J.M., and Schlüchter, C., 2010. Distribution,
> geometry, age and origin of overdeepened valleys and basins in the Alps and
> their foreland: Swiss Journal Geoscience, v. 103, p. 407–426.

> Sanders, J.W., Cuffey, K.M., Moore, J.R., MacGregor, K.R., and Kavanaugh,
> J.L., 2012. Periglacial weathering and headwall erosion in cirque glacier
> bergschrunds: Geology, v. 40, p. 779-782.

> Sternai, P., Herman, F., Valla, P.G., Champagnac, J.-D., 2013. Spatial and
> temporal variations of glacial erosion in the Rhône valley (Swiss Alps):
> Insights from numerical modeling. Earth and Planetary Science Letters, 368,
> p. 119-131.

> Valla, P.G., Shuster, D.L., van der Beek, P.A., 2011. Significant increase in
> relief of the European Alps during mid-Pleistocene glaciations. Nat. Geosci.
> 4, 688–692.