### 2021-05-16 articles/erosion/response-referee2.md
Add review from referee #2 Ian Evans.

<!-- Copyright (c) 2021, Julien Seguinot (juseg.github.io), Ian Delaney
Creative Commons Attribution-ShareAlike 4.0 International License
(CC BY-SA 4.0, http://creativecommons.org/licenses/by-sa/4.0/) -->

# Authors' response to Ian Evans

Dear Ian Evans,

Thank you very much for your detailed review of our manuscript.

## General Comments

> The authors provide large amounts of very useful results from simulations,
> concerning the effects of varying glacier geometry on erosion potential. It
> was quite an achievement to compute this at fairly high temporal and spatial
> resolution (1 km and 50 a throughout the Alps). The conclusions are
> reasonable, and I value the results. They provide support for the
> traditional view, that high-altitude features such as cirques are eroded by
> steep, local glaciers and not under extensive icefields well above ELA. Most
> erosion is expected along major valleys (glacial troughs): the potential
> erosion is greatest low down the glaciers, with a lesser maximum at high
> altitude and a minimum between. The paper is well written and the Figures
> are all relevant and interesting. The qualitative conclusions are well
> justified.

> An interesting result, not commented on, is that from Figs. 3 and 7 i & j,
> potential erosion volume in the Alps peaks when ice cover amounts to 1.6 cm
> sea level equivalent. I would love to see a map of the glaciers at that
> stage\*, which is well below maximum but considerably more than recent ice
> cover (0.3 ± 0.1 mm s.l.e., Farinotti et al. 2019). This peak applies to the
> ‘Koppes’ and ‘Herman’ erosion laws, but not to ‘Humphrey’ and ‘Cook’ laws
> which give very broad maxima for s.l.e. 10-30 cm (Fig. 7 k & l).

> As the comment (lines 221-222) on erosion distribution during advance,
> retreat and maxima is of great interest, it would be useful to illustrate
> this with three maps of potential erosion during each of these types of
> phase.

> I find no mention of lithology, rock type, geology, resistance or erodibility
> (excluding that citing Herman on line 56). This omission of half the erosion
> equation should be mentioned; it implies that the ‘potential’ nature of the
> erosion should be emphasised repeatedly. This adds to the admitted
> limitations of lack of feedback to changing topography, absence of hydrology
> and absence of subglacial sediment accumulation or movement. Using constant
> topography might be OK for the last glaciation, as here, but would become
> increasingly unrealistic for several glaciations. Destinations of eroded
> material are not considered: deposition in temporary storage may inhibit
> erosion in parts. The omission of hydrological processes must be a major
> drawback, as Herman et al. (2011), Egholm et al. (2012a, b) and Pedersen et
> al. (2014) showed that its inclusion of subglacial meltwater, plus horizontal
> stress gradients, gave more realistic simulations of erosion. For erosion
> rate as well as for cumulative erosion, it is potential erosion that is
> calculated, rather than predicted erosion.

> A further necessary geographical qualification concerns the misfit between
> simulated and actual glacier extent. The latter is well established from
> mapping of moraines. It is necessary to check Seguinot et al. (2018) to see
> that all the models predict too much ice in the southwest (south of 45° N)
> and far too much in the east (Sava, Drava, Mur, Enns and Traun valleys),
> while the EPICA models produce too little in the west (Rhône and Jura). I
> agree that this is probably because while the amount of precipitation is
> changed, the spatial pattern is not: thus the orographic effect of ice
> build-up increasing precipitation shadow, and the circulation changes due to
> southward shift of westerlies at LGM, are not included. To make this paper
> more self-contained, a brief mention of the 2018 evaluation is in order.
> Once a more realistic LGM precipitation pattern is accepted, future modelling
> might use gradations between that and the present pattern as a function of
> temperature.

> Overall, the paper reports on a very worthwhile computer-intensive exercise.
> It is densely packed with results and repays a careful read and a study of
> the videos. Clearly there can be interesting comparisons with Lai and Anders
> (2021) very recent submission to esurf !

>   [\*That stage can be approximated by pausing the first erosion
>    rates video around 99 ka and 13 ka. The three videos are strongly
>    recommended!]

> The text is short and pithy, so there is space for some extension. There are
> a few missed opportunities, and some incompleteness in the presentation. In
> detail I suggest the following improvements:

## Specific comments - Text:

> The 1 km resolution of the modelling is clearly important, so it is strange
> that the first mention is on p.11. Although it is very demanding of computer
> time, it is still (as admitted) not adequate to represent cirques (averaging
> of the order of 700 m across). Even stranger, there is no mention of the DEM
> used (it was SRTM in the 2018 paper). This information should be provided in
> section 2, p. 3-4.

> The lateral constraints on valley glaciers are usually covered by a ‘form
> factor’: how are they handled by this version of the PISM ice sheet model –
> how has lateral drag been included?

> It might be pointed out that the ‘Koppes’ relationship with an exponent of
> 2.34 is based on 13 data points that are very poorly distributed: in fact,
> Koppes et al. (2015) go on to drop two outliers and reach an exponent of
> 2.62.

> In lines 206-209, the point about trimlines, relating them to
> time-transgressive erosion, is well taken. But they represent the integrated
> effect of many phases of glacial erosion (as noted on 223-224), not
> necessarily just from recessional phases.

## - Figures:

> In Fig. 3 is almost impossible to follow the brown lines, unless advancing
> and retreating parts of lines are distinguished (e.g. by colour). (The ice
> volume video may help out.)  Also: keep to ‘potential erosion volume’. The
> pale brown bars show a decline in annual erosion volume with ice volumes over
> 1.8 cm s.l.e., stronger than is implied in line 110 of the text – it is more
> than ‘slight’.

> To aid interpretation of Figs 4 a-c and 7 a-d, a corresponding topographic
> map or hillshade of the Rhine area on the same scale should be provided.

> Fig. 1c illustrates a very unusual cirque, with a huge lake replacing a very
> unhealthy convex glacier. Fascinating, but surely a more representative
> cirque should be used. Actually these three photos are not well integrated
> with the text (pace lines 191 & 195-198). Lauterbrunnen is an extreme
> (vertical-sided) trough example, over-used by textbooks.

> Fig. 5: The elevation histogram in Fig. 5a could be misleading, as it seems
> to cover the whole study area, including parts never glacier-covered. It
> should be limited to areas glacier-covered at the maximum, and preferably
> supplemented by a histogram for those covered now or at the minimum. The
> accompanying main graph (5a) relates to areas glacier-covered at each time
> slice. Re another reviewer’s comment, yes LGM glaciers did reach below 500 m:
> at L. Garda (67 m?) and the Tagliamento Glacier lobe LGM ice reached below
> 100 m. An integrated plot of total-cycle potential erosion rate (and
> volume?) at each altitude, over glacier-covered areas, would be interesting
> (as an extra Figure): can rates at high altitude, where glacial duration is
> longer, catch up with those at low altitude?  Probably not, given the
> concentration of flow down major valleys. Striping above 3000 m, presumably
> related to low surface area, suggests a switch there to 20 m or 40 m bands
> would be advisable. Caption; ‘modelled potential erosion rates’?

> In Fig. 7, the quantitative contrasts between the ‘Koppes’ law (e) and the
> others are alarming. If the scales are taken literally, (e) gives potential
> erosion around 1 m, while (f) and (g) give hundreds of m, in 120 ka. As 120
> ka is only a fraction of the Quaternary, hundreds of m seems too much, while
> 1 m seems very low (0.008 mm a-1 overall). For me, there should be a correct
> answer between these extremes. The spatial pattern, however, is more
> important.

> Fig. 8a shows a flat dome at 3000 m for about the first 20 km of the Rhine.
> As there are mountains at 3630, 3583 and 3192 m around the sources, it is
> likely that   there would be steeper ice slopes up to these peaks, giving
> rather different stresses and erosion potentials. Also are the instabilities
> in Fig. 8b, near source and terminus, edge effects or related to the 1 km
> resolution?

>   The three videos are well worth watching  -in fact they clarify some
>   queries arising from the paper. Perhaps they could be captioned in the
>   ‘Video supplement’ paragraph. Some extra information on the third (bedrock
>   altitude) video could point out that there are numerous zero values not
>   shown: the zero geometric mean rates from 2300 to 3000 m around 24 ka
>   otherwise seem to conflict with the numerous positive rates plotted. Zero
>   rates presumably show where ice is frozen to the bed: their altitudinal
>   distribution might be worth discussion in the text. Above 3000 m means are
>   positive, which seems strange.

## Technical corrections:

> All map figures need km scale bars.

> Line 24  ‘are yet’

> 27 ‘in the absence’

> 38 ‘Examination … suggests’

> 61 ‘modelled erosion potential’ rather than ‘ modelled erosion rates’ ?

> 70 insert ‘isostatic’?

> 72 ‘and is run to…’

> 78 ‘temperature lapse-rate’ or is it ‘temperature change’ ?

> 116-7 ‘much of the’; and delete comma after bracket.

> 161 ‘does not increase’

> 190 delete both commas

## References not in original:

> Egholm, D.L., Pedersen, V.K., Knudsen, M.F., Larsen, N.K., 2012a. Coupling
> the flow of ice, water, and sediment in a glacial landscape evolution model.
> Geomorphology 141–142: 47–66.
> http://dx.doi.org/10.1016/j.geomorph.2011.12.019.

> Egholm, D.L., Pedersen, V.K., Knudsen, M.F., Larsen, N.K., 2012b. On the
> importance of higher-order ice dynamics for glacial landscape evolution.
> Geomorphology 141–142: 67-80.

> Farinotti, D., et al., 2019. A consensus estimate for the ice thickness
> distribution of all glaciers on Earth. Nat. Geosci. 12, 168–173.

> Lai, J., Anders, A.M., 2021. Climatic controls on mountain glacier basal
> thermal regimes dictate spatial patterns of glacial erosion.
> https://esurf.copernicus.org/preprints/esurf-2021-26/

> Pedersen, V.K., Huismans, R.S., Herman, F., Egholm, D.L., 2014. Controls of
> initial topography on temporal and spatial patterns of glacial erosion.
> Geomorphology 223, 96-116.

## Acknowledgements:

> I am grateful for helpful discussions with Iestyn Barr, Jeremy Ely, Matt
> Tomkins, Pippa Whitehouse, Cristina Balaban and Matt Wiecek.

> - Ian S. Evans, Durham University, U.K.